---
title: "Mammographic Masses Neural Network"
author: "Rex Manglicmot"
date: "2023-01-01"
output: 
  github_document: 
    toc: yes
---

## Continuing Working Document

## Introduction


## Loading the Libraries
```{r, message=FALSE}
library(tidyverse)
```


## Loading the Data
```{r}
#get data from UCI website
url <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/mammographic-masses/mammographic_masses.data'

#load data into object
data_orig <- read.csv(url)
```


## Cleaning the Data
```{r}
# look at data
str(data_orig)
```
It would seem that the data is messy.

On the UCI website it listed that the there are a total of 961 observatins but we have only 960. Therefore, the column name would appear to be an observation. However, upo further inspection, the "X5.1" and "X3.1" column does not make any sense because values should be based on 1-5 and 1-4, respectively. I will therefore delete that observation from the dataset and relabel the columns. 

But first, I will make a copy of the original dataset for manipulation.

```{r}
#make a copy
data <- data_orig

#change column names
colnames(data) <- c('BIRAD', 'age', 'shape', 'margin', 'density', 'result')

#get rid of BIRAD since on the UCI website it is non-predictive
data <- subset(data, select = -(BIRAD))

#How many NA do we have?
sum(is.na(data))
  
#how many columns have ?
sum(data == '?')

# which specific observations have '?'
which(data == '?')

# which specific observations have '*'
sum(data == '*')

#Replace ? with NAs
data[data == '?'] <- NA

#sum the NAs again
sum(is.na(data))

#remove the NA observations
data <- na.omit(data)

#dimensions of the dataset
dim(data)
```

Let's change the class
```{r}
#convert age column into a numeric
data$age <- as.numeric(as.character(data$age))

#create a list to convert remaining columns
index <- 2:ncol(data)

#pass list to function
data[ ,index] <- lapply(data[ ,index], as.factor)

#check structure again
str(data)
```
## Exploratory Data Analysis
```{r}
#see the summary statistics
summary(data)
```


## Neural Networks



## Limitations


## Conclusion


## Inspiration for this project